Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-02-19 analysis/paper/paper.Rmd
fix all typos

```{r sample-size}
# number of each phase
library(tidyverse)
pre_E <- burial_three_period_age_number %>% filter(Phase == "pre") %>% pull()
post_E <- burial_three_period_age_number %>% filter(Phase == "post") %>% pull()
chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
disturb <- burial_three_period_age_number %>% filter(Phase == "disturbed") %>% pull()
diff_pre_post <- post_E - pre_E
```

@@

## Hypothesis and construction of networks {-}
<!-- our expectation based on hypothesis --> 
Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000]. This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks. If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with associations with wealth accumulation, indicative of a network-based society. In contrast, the network before the European arrival will show low centralization and high transitivity, associated with ritual element, or age/sex groups to indicate a corporate-based society. To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan. By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.

@@

<!--posterior stats for pre-E-->
We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)). For nodal covariates in the pre-European model, the ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials. On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero. This demonstrates that burials in the pre-European period with ritual pottery tend to form relations. Despite positive mean values for some covariates, such as age and sex, they do not show a significant influence. Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment. For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect. The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common. In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties, without any prominent burials. This might imply that individuals have equal access to the flow of trade goods.

```{r table-posterior-stats}
# tidy data for the table of posterior stats 
posterior_esti_two_phases<-
  bergm_two_phases %>%
  pivot_longer(!phase, 
               names_to = "parameter", 
               values_to = "posterior") %>%
  mutate(parameter= case_when(parameter == "edges" ~ "Density",
                              parameter == "nodematch.age" ~ "Age-homophily",
                              parameter == "nodematch.sex" ~ "Sex-homophily",
                              parameter == "nodematch.ritual" ~ "Ritual-homophily",
                              parameter == "nodematch.value" ~ "Wealth-homophily",
                              parameter == "gwesp" ~ "Transitivity",
                              parameter == "gwdeg" ~ "Centralization",
                              parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
  mutate(parameter = factor(parameter,
                            levels=c("Density", "Age-homophily", 
                                     "Sex-homophily", "Ritual-homophily",
                                     "Wealth-homophily", "Transitivity",	
                                     "Centralization", "Physical distance"))) %>% 
  group_by(parameter, phase) %>% 
  summarize(mean = round(mean(posterior), 2),
            median = round(quantile(posterior, 0.5), 2),
            `2.5%` = round(quantile(posterior, 0.025), 2),
            `97.5%` = round(quantile(posterior, 0.975), 2)) %>% 
  pivot_wider(names_from = phase, 
              values_from = c(mean, median, `2.5%`, `97.5%`)) %>% 
  select(parameter, ends_with(c("pre", "post")))

library(flextable)
# add column labels
ft <- flextable(posterior_esti_two_phases) %>% 
  set_header_labels(mean_pre = "pre-European", median_pre = "pre-European", 
                    `2.5%_pre` = "pre-European", `97.5%_pre` = "pre-European",
                    mean_post = "post-European", median_post = "post-European", 
                    `2.5%_post` = "post-European", `97.5%_post` = "post-European") %>%
  merge_at(i = 1, j = 2:5, part = "header") %>% 
  merge_at(i = 1, j = 6:9, part = "header") %>% 
  add_header_row(values = c("", "mean", "median", "2.5%", "97.5%",
                            "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
  align(j = 2:9, align = "right", part = "header") %>% 
  align(i = 1, align = "center", part = "header") %>% 
  width(j = ~ parameter, width = 1.3) %>% 
  width(j = 2:9, width = 0.65) %>% 
  set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models. Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017). Posterior means are generally close to posterior medians.")
ft
```

```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases. The pre-European group presents remarkable larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group. The distance parameters overlap between two groups. Both posterior estimates present generally symmetric distributions."}
knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
```

<!--posterior stats for for post-E and their comparison-->
After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects. This indicates burials in the same wealth level tend to be more connected with each other. Despite negative effects for ritual and sex variables, they are not significant. Also, there are no significant effects for variables of age and physical distance. Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network. In contrast, centralization has a significantly higher positive effect than the pre-European network. This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others. This implies more access to trade goods and the increased wealth accumulation and display behaviors in burial events. In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model. This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.

@@

```{r figure-bootstrap-CI, fig.cap = "Results of the vertex bootstrap analysis by three network variables, showing the 95% bootstrap confidence intervals for the bootstrapped (n = 1000) networks, for each category of node removal. The 0% node removal category is the original data for the two phases. Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variables."}
knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
```

@@

<!--assessment of MCMC analysis and BGOF--> 
Our computational models reproduce networks that resemble the structural features of our observed networks. Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e. the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution. For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it. We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means. Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network. In general, this demonstrates a slightly better fit between model and data for the pre-European network.

@@

<!--archaeological implication based on the results-->
Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods. Our results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network. We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods. Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts. This might imply that ritual could be a status indicator at the time before foreign goods were first introduced to Indigenous societies. After the European presence was established, status indicators shifted from ritual to wealth differences. Moreover, geographical proximity, implying some kinship association, has not effect on sharing similar prestige items between burials for both phases. This may suggest that the use of prestige items is more individual-oriented instead of kinship-based. Demographic information, such as age and sex, do not have any effects on network formation for both networks. But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.

@@

<!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
We examined social changes in a European pericolonial context in northeastern Taiwan. We tested the hypotheses of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects. The results support our model that the relationship between burials changed after the European colonization period in the 17th century. Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism. After European arrival, the network has a tendency of centralization relative to the rarity of goods. We interpret the changes in formation mechanisms of networks, from ritual practice to wealth levels, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network. This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000]. Using burial data with historical documents, we are to better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g. @Acabado2017; @Trabert2017].

>>>>

```{r sample-size}
# number of each phase
library(tidyverse)
pre_E <- burial_three_period_age_number %>% filter(Phase == "pre") %>% pull()
post_E <- burial_three_period_age_number %>% filter(Phase == "post") %>% pull()
chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
disturb <- burial_three_period_age_number %>% filter(Phase == "disturbed") %>% pull()
diff_pre_post <- post_E - pre_E
pre_and_post <- pre_E + post_E

# missing values
age_miss <- sum(NAs_age$n)
sex_miss <- sum(NAs_sex$n)
```

@@

## Hypothesis and construction of networks {-}

<!-- our expectation based on hypothesis --> 
Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000]. This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks. If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with associations with wealth accumulation, indicative of a network-based society. In contrast, the network before the European arrival will show low centralization and high transitivity, associated with ritual element, or age/sex groups to indicate a corporate-based society. To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan. By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.

@@

<!--posterior stats for pre-E-->
We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)). For nodal covariates in the pre-European model, the ritual element represented by pots (ritual-homophily) has a significant effect on the formation of relations between burials. On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero. This demonstrates that burials in the pre-European period with ritual pottery tend to form relations. Despite positive mean values for some covariates, such as age and sex, they do not show a significant influence. Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment. For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect. The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common. In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties, without any prominent burials. This might imply that individuals have equal access to the flow of trade goods.

```{r table-posterior-stats}
# tidy data for the table of posterior stats 
posterior_esti_two_phases<-
  bergm_two_phases %>%
  pivot_longer(!phase, 
               names_to = "parameter", 
               values_to = "posterior") %>%
  mutate(parameter= case_when(parameter == "edges" ~ "density",
                              parameter == "nodematch.age" ~ "age-homophily",
                              parameter == "nodematch.sex" ~ "sex-homophily",
                              parameter == "nodematch.ritual" ~ "ritual-homophily",
                              parameter == "nodematch.value" ~ "wealth-homophily",
                              parameter == "gwesp" ~ "transitivity",
                              parameter == "gwdeg" ~ "centralization",
                              parameter == "dyadcov.distance" ~ "physical distance")) %>% 
  mutate(parameter = factor(parameter,
                            levels=c("density", "age-homophily", 
                                     "sex-homophily", "ritual-homophily",
                                     "wealth-homophily", "transitivity",	
                                     "centralization", "physical distance"))) %>% 
  group_by(parameter, phase) %>% 
  summarize(mean = round(mean(posterior), 2),
            median = round(quantile(posterior, 0.5), 2),
            `2.5%` = round(quantile(posterior, 0.025), 2),
            `97.5%` = round(quantile(posterior, 0.975), 2)) %>% 
  pivot_wider(names_from = phase, 
              values_from = c(mean, median, `2.5%`, `97.5%`)) %>% 
  select(parameter, ends_with(c("pre", "post")))

library(flextable)
# add column labels
ft <- flextable(posterior_esti_two_phases) %>% 
  colformat_double(digits = 2) %>% 
  set_header_labels(mean_pre = "pre-European", median_pre = "pre-European", 
                    `2.5%_pre` = "pre-European", `97.5%_pre` = "pre-European",
                    mean_post = "post-European", median_post = "post-European", 
                    `2.5%_post` = "post-European", `97.5%_post` = "post-European") %>%
  merge_at(i = 1, j = 2:5, part = "header") %>% 
  merge_at(i = 1, j = 6:9, part = "header") %>% 
  add_header_row(values = c("", "mean", "median", "2.5%", "97.5%",
                            "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
  align(j = 2:9, align = "right", part = "header") %>% 
  align(i = 1, align = "center", part = "header") %>% 
  width(j = ~ parameter, width = 1.36) %>% 
  width(j = 2:9, width = 0.67) %>% 
  set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models. Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017). Posterior means are generally close to posterior medians.")
ft
```

```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases. The pre-European group presents remarkably larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group. The distance parameters overlap between two groups. Both posterior estimates present generally symmetric distributions."}
knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
```

<!--posterior stats for for post-E and their comparison-->
After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects. This indicates burials in the same wealth level tend to be more connected with each other. Despite negative effects for ritual and sex variables, they are not significant. Also, there are no significant effects for variables of age and physical distance. Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network. In contrast, centralization has a significantly higher positive effect than the pre-European network. This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others. This implies more access to trade goods, increased wealth accumulation and display behaviors in burial events. In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model. This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.

@@

```{r figure-bootstrap-CI, fig.cap = "Results of the vertex bootstrap analysis by three network variables, showing the 95% bootstrap confidence intervals for the bootstrapped (n = 1000) networks, for each category of node removal. The 0% node removal category is the original data for the two phases. Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variable."}
knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
```

@@

<!--assessment of MCMC analysis and BGOF--> 
Our computational models reproduce networks that resemble the structural features of our observed networks. Bayesian goodness of fit diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases); details in Supplementary Online Materials) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e. the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution. For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it. We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means. Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network. In general, this demonstrates a slightly better fit between model and data for the pre-European network.

@@

<!--archaeological implication based on the results-->
Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods. Our results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network. We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods. Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts. This might imply that ritual could be a status indicator at the time before many foreign goods were introduced to Indigenous societies. After the European presence was established, status indicators shifted from ritual to wealth differences. Moreover, geographical proximity, implying some kinship association, has no effect on sharing similar prestige items between burials for both phases. This may suggest that the use of prestige items is more individual-oriented instead of kinship-based. Demographic information, such as age and sex, do not have any effects on network formation for both networks. But we also note that no effects could be due to the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.

@@

<!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
We examined social changes in a European pericolonial context in northeastern Taiwan. We tested the hypotheses of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects. The results support our model that the relationship between burials changed after the European colonization period in the 17th century. Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism. After European arrival, the network has a tendency of centralization relative to the rarity of goods. We interpret the changes in formation mechanisms of networks, from ritual practice to wealth levels, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network. This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000]. Using burial data with historical documents, we can better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g. @Acabado2017; @Trabert2017].

### 2021-02-19 analysis/paper/supplementary-material.Rmd
fix all typos

# Kiwulan in the 17th northstern Taiwan in a pericolonial context {-}

<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch. The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016]. In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements. The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].

@@

Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials. Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015]. Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].

@@

<!-- burial data and the definition of tie formation: partially move to SI  -->
For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)). Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For burial attributes, we assigned burials to different groups according to their age, sex, burial value and the presence/absence of ritual pottery. We have three groups for the age attribute, 0-12, 12-20, and above 20; two groups for the sex attribute, male and female, and four groups for the burial value that are specified in the main paper document.

@@

@@

<!--Model fitting: move to SI -->
After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on he approximate exchange algorithm [@Caimo2011]. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pairs of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`. We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.

>>>>

# Kiwulan in the 17th northeastern Taiwan in a pericolonial context {-}

<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch. The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016]. In the annual event held by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan sticks or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements. The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].

@@

Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials. Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015]. Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Gjesfjeld2015].

@@

<!-- burial data and the definition of tie formation: partially move to SI  -->
For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)). Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For burial attributes, we assigned burials to different groups according to their age, sex, burial value and the presence/absence of ritual pottery. We have three groups for the age attribute, 0-12, 12-20, and above 20; two groups for the sex attribute, male and female; and four groups for the burial value that are specified in the main paper document. Table \@ref(tab:age-sex-table) shows the number of burials, of which sex and age can be determined.

@@

```{r age-sex-table}
library(flextable)
age_sex_by_phase <-
  burial_three_period_age_tidy %>% 
  select(Age_scale, gender, Phase) %>% 
  group_by(Phase, gender, Age_scale) %>% 
  filter(!Phase %in% c("disturbed", "chi")) %>% 
  count() %>% 
  pivot_wider(names_from = c(Phase, gender), 
              values_from = n) %>% 
  mutate(Age_scale = factor(Age_scale,
                            levels = c("0-12", "12-20", "+20", "NA"))) %>% 
  arrange(Age_scale) %>% 
  rename(age = Age_scale) %>% 
  select(age, pre_male, pre_female, pre_NA, post_male, post_female, post_NA)

ft_age_sex <- flextable(age_sex_by_phase) %>% 
  set_header_labels(pre_male = "pre-European", pre_female = "pre-European",
                    pre_NA = "pre-European", post_male = "post-European", 
                    post_female ="post-European", post_NA = "post-European") %>%
  merge_at(i = 1, j = 2:4, part = "header") %>% 
  merge_at(i = 1, j = 5:7, part = "header") %>% 
  add_header_row(values = c("","male", "female", "NA", 
                            "male", "female", "NA"), top = FALSE) %>% 
  align(j = 2:7, align = "right", part = "header") %>% 
  align(i = 1, align = "center", part = "header") %>% 
  set_caption("The number of burials with identification of age and sex at Kiwulan. The NA means no sufficient skeletal remains for determination of age or sex")
ft_age_sex
```

@@

<!--Model fitting: move to SI -->
After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on the approximate exchange algorithm [@Caimo2011]. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pairs of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`. We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.
